On the specific recursive structure of iterated parametricity

### 2025-11-09 paper/paper.tex
paper: near functional completion

\newcommandx{\deps}[2][1,2]{\prim{deps}[][#1][#2][][]}
\newcommandx{\depscohs}[2][1,2]{\prim{depscohs}[][#1][#2][][]}
\newcommandx{\depstype}[2][1,2]{\prim{depstype}[][#1][#2][][]}
\newcommandx{\depscohstype}[2][1,2]{\prim{depscohstype}[][#1][#2][][]}

@@

\renewcommandx{\framep}[2][1,2]{\prim{frame}[][#1][#2][][]}
\renewcommandx{\layer}[2][1,2]{\prim{layer}[][#1][#2][][]}
\renewcommandx{\painting}[2][1,2]{\prim{painting}[][#1][#2][][]}
\renewcommandx{\restrf}[4][1,2,3,4]{\restr{frame}[][#3][#4][#1][#2][][]}
\renewcommandx{\restrl}[4][1,2,3,4]{\restr{layer}[][#3][#4][#1][#2][][]}
\renewcommandx{\restrp}[4][1,2,3,4]{\restr{painting}[][#3][#4][#1][#4][][]}
\renewcommandx{\cohf}[6][1,2,3,4,5,6]{\coh{frame}[][#3][#4][#5][#6][#1,#2][][]}
\renewcommandx{\cohl}[6][1,2,3,4,5,6]{\coh{layer}[][#3][#4][#5][#6][#1,#2][][]}
\renewcommandx{\cohp}[6][1,2,3,4,5,6]{\coh{painting}[][#3][#4][#5][#6][#1,#2][][]}

@@

              \end{array}\right) : \frametype[n][0,...,p]

@@

\begin{equation*}
  \fontsize{7.2}{9}\selectfont
  \begin{array}{llcl}
    \painting[n][0,...,p]          &
    \multicolumn{3}{l}{
      \left(\begin{array}{l}
                \deps[n-1][0,...,n-1]:\depstype[n-1][0,...,n-1] \\
                E:\framep[n-1][n-1]\rightarrow \U               \\
              \end{array}\right) : \paintingtype[n][0,...,p](\framep[n][0,...,p](\deps[n-1][0,...,p]))
    }                                                            \\
    \painting[n][0,...,0-1]          &
    \left(\begin{array}{l}
              \unitpoint \\
              E                     \\
            \end{array}\right) & \defeq & \unittype                        \\
    \painting[n][0,...,p] &
    \left(\begin{array}{l}
              \deps[n-1][0,...,p-1] \\
              E                     \\
            \end{array}\right) & \defeq &
       (\painting[n][0,...,p-1]
       \left(\begin{array}{l}
           \deps[n-1][0,...,n-1] \\
           E                     \\
         \end{array}\right),           \\
       \painting[n][p]
       \left(\begin{array}{l}
           \deps[n-1][0,...,n-1] \\
           E                     \\
         \end{array}\right))           \\
  \end{array}
\end{equation*}

@@

        \deps.layer[n][p](\deps[n][0,...,n])(\restrf[n][p]

@@

\begin{align*}
  \overline{\deps}(\depscohs:\depscohstype) : \depstype\\
  \overline{\deps}(\depscohs) \defeq
\left(\begin{array}{l}
            \framep[n][p](\depscohs.\deps[n][0,...,p])   \\
            \painting[n][p](\depscohs.\deps[n][0,...,n]) \\
            \restrf[n][p](\depscohs[n][0,...,p])\\
      \end{array}\right)\\
\end{align*}

Armed with $\restrframe$ and $\cohftype$, it is now possible to define
$\restrp[n][p]$ by recursion on $n-p$:

\begin{equation*}
  \fontsize{7.2}{9}\selectfont
  \begin{array}{llcl}
    \restrp[n][p]          &
    \multicolumn{3}{l}{
      \left(\begin{array}{l}
                \depscohs[n][0,...,n]:\depscohstype[n][0,...,n] \\
                E:\framep[n-1][n-1]\rightarrow \U               \\
                E':\framep[n][n]\rightarrow \U               \\
      \end{array}\right) & : &
      \Pi q\leq n. \Pi \epsilon. \Pi d:\framep[n][p]
      \left(\begin{array}{l}
                \overline{\deps}(\depscohs)\\
                E':\framep[n][n]\rightarrow \U               \\
      \end{array}\right).\,
      \painting[n][p]
      \left(\begin{array}{l}
                \overline{\deps}(\depscohs)\\
                E':\framep[n][n]\rightarrow \U               \\
      \end{array}\right)(d) \rightarrow
      \painting[n-1][p]
      \left(\begin{array}{l}
                \depscohs[n][0,...,n].\deps\\
                E:\framep[n-1][n-1]\rightarrow \U               \\
      \end{array}\right)(\restrf[n][p](\depscohs[n][0,...,n].deps)(d))
    }                                                            \\
    \restrp[n][n]          &
    \left(\begin{array}{l}
              \depscohs[n][0,...,n] \\
              E                     \\
              E'                    \\
            \end{array}\right) & \defeq & \lambda q\leq n.\lambda \epsilon. \lambda d. \lambda (l,c). l\epsilon                     \\
    \restrp[n][p](p<n)     &
    \left(\begin{array}{l}
              \depscohs[n][0,...,n] \\
              E                     \\
              E'                    \\
            \end{array}\right) & \defeq & \lambda q\leq n.\lambda \epsilon. \lambda d. \lambda (l,c).
    \left(\begin{array}{l}
              \restrl[n][p][q][\epsilon](\depscohs[n-1][0,...,p-1])(l), \\
              \restrp[n][p][q][\epsilon]
              \left(\begin{array}{l}
                  \depscohs[n][0,...,n] \\
                  E                     \\
                  E'                    \\
                \end{array}\right)(\restrf[n][p](\depscohs[n][0,...,p])(d))(c)           \\
            \end{array}\right) \\
  \end{array}
\end{equation*}

@@

\begin{equation*}
  \fontsize{7.2}{9}\selectfont
  \begin{array}{llcl}
    \cohptype[n][0,...,p]                    & (\depscohs[n][0,...,n]:\depscohstype[n][0,...,n]) & :      & \U                                        \\
    \cohptype[n][]                           & \unitpoint                                & \defeq & \unittype                                 \\
    \cohptype[n][0,...,p](\depscohs[n][0,...,n]) & (\depscohs[n][0,...,n])                       & \defeq & \cohptype[n+1][0,...,p-1] \times \left(
    \begin{array}{l}
        \Pi q\leq n.\Pi \epsilon.\Pi d:\framep(\depscohs[n][0,...,p]). \\
        \painting[n][p](\depscohs[n][0,...,n])(d) \rightarrow          \\ \depscohs.\painting[n][p](\depscohs.\restrf[n][p][q][\epsilon](d))\end{array}
    \right)                                                                                                                                     \\
  \end{array}
\end{equation*}

@@

              \end{array}\right) : \cohftype[n][0,...,p]

@@

      (\cohf[n+1][0,...,p], \lambda q.\lambda r.\lambda \epsilon.\lambda \omega.\lambda (d,l). \\
      \left(\cohf[n][0,...,p][q][r][\epsilon][\omega]

@@

                \end{array}\right) (d)(l)\right)                                                                           \\

@@

      \end{array}\right)\\

@@

    \lambda d. \lambda l. \lambda \theta. (\cohp[n][p][q][\epsilon][r][\omega](l))                                                                    \\

@@

\bibliographystyle{msclike}
\bibliography{paper}

>>>>

\newcommandx{\deps}[2][1,2]{\prim{deps}[][#1][#2][][]}
\newcommandx{\depscohs}[2][1,2]{\prim{depscohs}[][#1][#2][][]}
\newcommandx{\depscoht}[2][1,2]{\prim{depscoht}[][#1][#2][][]}
\newcommandx{\depstype}[2][1,2]{\prim{depstype}[][#1][#2][][]}
\newcommandx{\depscohstype}[2][1,2]{\prim{depscohstype}[][#1][#2][][]}
\newcommandx{\depscohttype}[2][1,2]{\prim{depscohttype}[][#1][#2][][]}

@@

\renewcommandx{\framep}[2][1,2]{\prim{frame}[][#1][#2][][]}
\renewcommandx{\layer}[2][1,2]{\prim{layer}[][#1][#2][][]}
\renewcommandx{\painting}[2][1,2]{\prim{painting}[][#1][#2][][]}
\renewcommandx{\restrf}[4][1,2,3,4]{\restr{frame}[][#3][#4][#1][#2][][]}
\renewcommandx{\restrl}[4][1,2,3,4]{\restr{layer}[][#3][#4][#1][#2][][]}
\renewcommandx{\restrp}[4][1,2,3,4]{\restr{painting}[][#3][#4][#1][#2][][]}
\renewcommandx{\cohf}[6][1,2,3,4,5,6]{\coh{frame}[][#3][#4][#5][#6][#1,#2][][]}
\renewcommandx{\cohl}[6][1,2,3,4,5,6]{\coh{layer}[][#3][#4][#5][#6][#1,#2][][]}
\renewcommandx{\cohp}[6][1,2,3,4,5,6]{\coh{painting}[][#3][#4][#5][#6][#1,#2][][]}

@@

            \end{array}\right) & :      & \frametype[n][0,...,p]                             \\

@@

\begin{equation*}
  \fontsize{7.2}{9}\selectfont
  \begin{array}{llcl}
    \painting[n][0,...,p]                              &
    \left(\begin{array}{l}
              \deps[n-1][0,...,n-1]:\depstype[n-1][0,...,n-1] \\
              E:\framep[n-1][n-1]\rightarrow \U               \\
            \end{array}\right) & :      & \paintingtype[n][0,...,p](\framep[n][0,...,p](\deps[n-1][0,...,p])) \\
    \painting[n][0,...,0-1]                            &
    \left(\begin{array}{l}
              \unitpoint \\
              E          \\
            \end{array}\right)                             & \defeq & \unittype                               \\
    \painting[n][0,...,p]                              &
    \begin{array}{l}
      \deps[n-1][0,...,p-1] \\
      E                     \\
    \end{array}                           & \defeq &
    \left(\painting[n][0,...,p-1]
    \left(\begin{array}{l}
                \deps[n-1][0,...,n-1] \\
                E                     \\
              \end{array}\right),\,
    \painting[n][p]
    \left(\begin{array}{l}
                \deps[n-1][0,...,n-1] \\
                E                     \\
              \end{array}\right)\right)                                                                         \\
  \end{array}
\end{equation*}

@@

        \deps.\layer[n][p](\deps[n][0,...,n])(\restrf[n][p]

@@

\begin{align*}
  \overline{\deps}(\depscohs:\depscohstype) : \depstype \\
  \overline{\deps}(\depscohs) \defeq
  \left(\begin{array}{l}
            \framep[n][p](\depscohs.\deps[n][0,...,p])   \\
            \painting[n][p](\depscohs.\deps[n][0,...,n]) \\
            \restrf[n][p](\depscohs[n][0,...,p])         \\
          \end{array}\right)    \\
\end{align*}

Armed with $\restrf$ and $\cohftype$, it is now possible to define
$\restrp[n][p]$ by recursion on $n-p$:

\begin{equation*}
  \fontsize{7.2}{9}\selectfont
  \begin{array}{llcl}
    \restrp[n][p]                                      &
    \left(\begin{array}{l}
              \depscohs[n][0,...,n]:\depscohstype[n][0,...,n] \\
              E:\framep[n-1][n-1]\rightarrow \U               \\
              E':\framep[n][n]\rightarrow \U                  \\
            \end{array}\right) & :      &
    \left(
    \begin{array}{l}
        \Pi q\leq n. \Pi \epsilon. \Pi d:\framep[n][p]
        \left(\begin{array}{l}
                  \overline{\deps}(\depscohs)    \\
                  E':\framep[n][n]\rightarrow \U \\
                \end{array}\right).    \\
        \painting[n][p]
        \left(\begin{array}{l}
                  \overline{\deps}(\depscohs)    \\
                  E':\framep[n][n]\rightarrow \U \\
                \end{array}\right)(d) \rightarrow \\
        \painting[n-1][p]
        \left(\begin{array}{l}
                  \depscohs[n][0,...,n].\deps       \\
                  E:\framep[n-1][n-1]\rightarrow \U \\
                \end{array}\right)(\restrf[n][p](\depscohs[n][0,...,n].deps)(d))\end{array}\right)                                                  \\
    \restrp[n][n]                                      &
    \left(\begin{array}{l}
              \depscohs[n][0,...,n] \\
              E                     \\
              E'                    \\
            \end{array}\right)                           & \defeq & \lambda q\leq n.\lambda \epsilon. \lambda d. \lambda (l,c). l\epsilon \\
    \restrp[n][p](p<n)                                 &
    \left(\begin{array}{l}
              \depscohs[n][0,...,n] \\
              E                     \\
              E'                    \\
            \end{array}\right)                           & \defeq & \lambda q\leq n.\lambda \epsilon. \lambda d. \lambda (l,c).
    \left(\begin{array}{l}
              \restrl[n][p][q][\epsilon](\depscohs[n-1][0,...,p-1])(l),            \\
              \restrp[n][p][q][\epsilon]
              \left(\begin{array}{l}
                  \depscohs[n][0,...,n] \\
                  E                     \\
                  E'                    \\
                \end{array}\right)(\restrf[n][p](\depscohs[n][0,...,p])(d))(c) \\
            \end{array}\right)                                                          \\
  \end{array}
\end{equation*}

@@

\begin{equation*}
  \fontsize{7.2}{9}\selectfont
  \begin{array}{llcl}
    \cohptype[n][0,...,p] & (\depscohs[n][0,...,n]:\depscohstype[n][0,...,n]) & :      & \U                                      \\
    \cohptype[n][]        & \unitpoint                                        & \defeq & \unittype                               \\
    \cohptype[n][0,...,p] & (\depscohs[n][0,...,n])                           & \defeq & \cohptype[n+1][0,...,p-1] \times \left(
    \begin{array}{l}
        \Pi q\leq n.\Pi r\leq q.\Pi \epsilon.\Pi \omega.\Pi d:\framep(\depscohs[n][0,...,p]). \\
        \painting[n][p](\depscohs[n][0,...,n])(d) \rightarrow                                 \\ \depscohs.\painting[n][p](\depscohs.\restrf[n][p][q][\epsilon](d))\end{array}
    \right)                                                                                                                      \\
  \end{array}
\end{equation*}

@@

            \end{array}\right)          & :      & \cohftype[n][0,...,p]                                                                                                     \\

@@

              \lambda q.\lambda r.\lambda \epsilon.\lambda \omega.\lambda (d,l). \\
              \cohf[n][0,...,p][q][r][\epsilon][\omega]

@@

                \end{array}\right)(d)(l)                                     \\

@@

@@

    \lambda d. \lambda l. \lambda \theta. (\cohp[n][p][q][\epsilon][r][\omega](l))  ...                                                                                    \\

@@

This is time for a new abbreviation:

\begin{align*}
  \depscohttype[n][0,..,p-1] \defeq
  \left(\begin{array}{l}
            \depscohs[n][0,...,n] :\depscohstype[n][0,...,n] \\
            \cohp[n][0,...p-1]: \cohptype[n][0,...,p-1]      \\
            \coht[n][0,...,p-1]: \cohttype[n][0,...,p-1]     \\
          \end{array}\right)
\end{align*}

The next step is to define $\cohp[n][p]$ by recursion on $n-p$:

\begin{equation*}
  \fontsize{7.2}{9}\selectfont
  \begin{array}{llcl}
    \cohp[n][p]                                        &
    \left(\begin{array}{l}
              \depscoht[n][0,...,n]:\depscohttype[n][0,...,n] \\
              E:\framep[n-2][n-2]\rightarrow \U               \\
              E':\framep[n-1][n-1]\rightarrow \U              \\
              E'':\framep[n][n]\rightarrow \U                 \\
            \end{array}\right) & :      &
    \left(
    \begin{array}{l}
        \Pi q\leq n. \Pi \epsilon. \Pi d:\framep[n][p]
        \left(\begin{array}{l}
                  \overline{\depscohs}(\depscoht) \\
                  E':\framep[n][n]\rightarrow \U  \\
                \end{array}\right).   \\
        \painting[n][p]
        \left(\begin{array}{l}
                  \overline{\depscohs}(\depscoht) \\
                  E':\framep[n][n]\rightarrow \U  \\
                \end{array}\right)(d) \rightarrow \\
        \painting[n-1][p]
        \left(\begin{array}{l}
                  \depscoht[n][0,...,n].\deps       \\
                  E:\framep[n-1][n-1]\rightarrow \U \\
                \end{array}\right)(\cohf[n][p](\depscoht[n][0,...,n].deps)(d))\end{array}\right)                                                  \\
    \cohp[n][n]                                        &
    \left(\begin{array}{l}
              \depscoht[n][0,...,n] \\
              E                     \\
              E'                    \\
            \end{array}\right)                           & \defeq & \lambda q\leq n.\lambda \epsilon. \lambda d. \lambda (l,c). l\epsilon \\
    \cohp[n][p](p<n)                                   &
    \left(\begin{array}{l}
              \depscoht[n][0,...,n] \\
              E                     \\
              E'                    \\
            \end{array}\right)                           & \defeq & \lambda q\leq n.\lambda \epsilon. \lambda d. \lambda (l,c).
    \left(\begin{array}{l}
              \cohl[n][p][q][\epsilon](\depscoht[n-1][0,...,p-1])(l),            \\
              \cohp[n][p][q][\epsilon]
              \left(\begin{array}{l}
                  \depscoht[n][0,...,n] \\
                  E                     \\
                  E'                    \\
                \end{array}\right)(\cohf[n][p](\depscoht[n][0,...,p])(d))(c) \\
            \end{array}\right)                                                           \\
  \end{array}
\end{equation*}

From which we can deduce $\cohp[n][0,...,p]$ again by a second recursion, this time on $p$:

Armed with all the constructions we can now define
\bibliographystyle{msclike}
\bibliography{paper}
